---
title: "Example 1"
output:
  html_document:
    df_print: paged
---

# Load functions
```{r}
source("evaluate_samples.R")
source("generate_samples.R")
library(ggplot2)
library(dplyr)
```

# Example 1: parallel design
In the examples there are 2 groups: treatment y control
```{r}
n_total <- 10000
n <- n_total/2
# Groups
group_1 <- c(rep(TRUE, n), rep(FALSE, n))
group_2 <- !group_1
```


## 1.A no treatment effect
First, generate the data with no treatment effect.
```{r}
set.seed(123)
# Generate data
Y_1a <- generate.samples.parallel(
  n_total=n_total, b0=120, sigma_u=20,
  treatment_effect=0
  )
# Basic summary
mean_1a <- mean(Y_1a)
sd_1a <- sd(Y_1a)
sd_1a_group1 <- sd(Y_1a[group_1])
sd_1a_group2 <- sd(Y_1a[group_2])
cat(mean_1a)
cat(sd_1a, sd_1a_group1, sd_1a_group2)
```

Plot the distribution of both groups
```{r}
df_1a <- data.frame(
  values=Y_1a, groups=c(rep("Group 1", n), rep("Group 2", n)))
ggplot(df_1a) +
  geom_density(aes(x=values, fill=groups), alpha=0.4) +
  theme_classic()
```

## 1.B yes treatment effect
```{r}
set.seed(123)
# Generate data
Y_1b <- generate.samples.parallel(
  n_total=n_total, b0=120, sigma_u=20,
  treatment_effect=0.4
  )
# Basic summary
mean_1b <- mean(Y_1b)
mean_1b_group1 <- mean(Y_1b[group_1])
mean_1b_group2 <- mean(Y_1b[group_2])
sd_1b <- sd(Y_1b)
sd_1b_group1 <- sd(Y_1b[group_1])
sd_1b_group2 <- sd(Y_1b[group_2])
cat(mean_1b, mean_1b_group1, mean_1b_group2)
cat(sd_1b, sd_1b_group1, sd_1b_group2)
# Plot
df_1b <- data.frame(
  values=Y_1b, groups=c(rep("Group 1", n), rep("Group 2", n)))
g <- ggplot(df_1b) +
  geom_density(aes(x=values, fill=groups), alpha=0.4) +
  theme_classic()
g
```



# Example 2: crossover design

Find the groups for the next 2 cases
```{r}
# Find groups
# Sequence in arrays: AB_P1, BA_P1, AB_P2, BA_P2
n_total <- 10000
n <- n_total/4
group_AB_P1 <- c(rep(TRUE, n), rep(FALSE, 3*n))
group_BA_P1 <- c(rep(FALSE, n), rep(TRUE, n), rep(FALSE, 2*n))
group_AB_P2 <- c(rep(FALSE, 2*n), rep(TRUE, n), rep(FALSE, n))
group_BA_P2 <- c(rep(FALSE, 3*n), rep(TRUE, n))
group_AB <- c(group_AB_P1 | group_AB_P2)
group_BA <- !group_AB
group_treatment <- c(group_AB_P2 | group_BA_P1)
group_control <- !group_treatment
group_period1 <- c(group_AB_P1 | group_BA_P1)
group_period2 <- !group_period1
```

## 2.A No treatment or period effects
```{r}
set.seed(123)
# Generate data
Y_2a <- generate.samples.crossover(
  n_total=n_total, b0=120, sigma_u=20, rho=0.75, # for 2x between var
  treatment_effect=0, period_effect=0
  )
# Basic summary
mean_2a <- mean(Y_2a)
mean_AB <- mean(Y_2a[group_AB])
mean_BA <- mean(Y_2a[group_BA])
sd_2a <- sd(Y_2a)
sd_AB <- sd(Y_2a[group_AB])
sd_BA <- sd(Y_2a[group_BA])
sd_within <- sqrt(get.var.within.crossover(Y_2a))
cat(mean_2a, mean_AB, mean_BA)
cat(sd_2a, sd_AB, sd_BA, sd_within)
# Plot
df_2a <- data.frame(
  values=Y_2a, group_vector=group_AB) %>%
  mutate(groups=if_else(group_AB, "AB", "BA"))
ggplot(df_2a) +
  geom_density(aes(x=values, fill=groups), alpha=0.4) +
  theme_classic()
```



## 2.B Only treatment effects

```{r}
set.seed(123)
# Generate data
Y_2b <- generate.samples.crossover(
  n_total=n_total, b0=120, sigma_u=20, rho=0.75,
  treatment_effect=0.4, period_effect=0
  )
# Basic summary
# Means
mean_2b <- mean(Y_2b)
mean_AB <- mean(Y_2b[group_AB])
mean_BA <- mean(Y_2b[group_BA])
mean_treatment <- mean(Y_2b[group_treatment])
mean_control <- mean(Y_2b[group_control])
mean_P1 <- mean(Y_2b[group_period1])
mean_P2 <- mean(Y_2b[group_period2])
# Sds
sd_2b <- sd(Y_2b)
sd_AB <- sd(Y_2b[group_AB])
sd_BA <- sd(Y_2b[group_BA])
sd_treatment <- sd(Y_2b[group_treatment])
sd_control <- sd(Y_2b[group_control])
sd_P1 <- sd(Y_2b[group_period1])
sd_P2 <- sd(Y_2b[group_period2])
sd_within <- sqrt(get.var.within.crossover(Y_2b))
# Print
cat(mean_2b, mean_AB, mean_BA)
cat(mean_treatment, mean_control, mean_P1, mean_P2)
cat(sd_2b, sd_AB, sd_BA, sd_within)
cat(sd_treatment, sd_control, sd_P1, sd_P2)
# Plot group
df_2b <- data.frame(values=Y_2b) %>%
  mutate(groups=if_else(group_AB, "AB", "BA"),
         treatment=if_else(group_treatment, "Treatment", "Control"),
         period=if_else(group_period1, "P1", "P2"),
         eachfour=c(rep("AB_P1", n), rep("BA_P1", n),
                    rep("AB_P2", n), rep("BA_P2", n))
  )
ggplot(df_2b) +
  geom_density(aes(x=values, fill=groups), alpha=0.4) +
  theme_classic()
ggplot(df_2b) +
  geom_density(aes(x=values, fill=treatment), alpha=0.4) +
  theme_classic()
ggplot(df_2b) +
  geom_density(aes(x=values, fill=period), alpha=0.4) +
  theme_classic()
ggplot(df_2b) +
  geom_density(aes(x=values, fill=eachfour), alpha=0.4) +
  theme_classic()
```



## 2.C Only period effects
```{r}
set.seed(123)
# Generate data
Y_2c <- generate.samples.crossover(
  n_total=n_total, b0=120, sigma_u=20, rho=0.75,
  treatment_effect=0, period_effect=0.03
  )
# Basic summary
# Means
mean_2c <- mean(Y_2c)
mean_AB <- mean(Y_2c[group_AB])
mean_BA <- mean(Y_2c[group_BA])
mean_treatment <- mean(Y_2c[group_treatment])
mean_control <- mean(Y_2c[group_control])
mean_P1 <- mean(Y_2c[group_period1])
mean_P2 <- mean(Y_2c[group_period2])
# Sds
sd_2c <- sd(Y_2c)
sd_within <- sqrt(get.var.within.crossover(Y_2c))
sd_AB <- sd(Y_2c[group_AB])
sd_BA <- sd(Y_2c[group_BA])
sd_treatment <- sd(Y_2c[group_treatment])
sd_control <- sd(Y_2c[group_control])
sd_P1 <- sd(Y_2c[group_period1])
sd_P2 <- sd(Y_2c[group_period2])
# Print
cat(mean_2c, mean_AB, mean_BA)
cat(mean_treatment, mean_control, mean_P1, mean_P2)
cat(sd_2c, sd_AB, sd_BA, sd_within)
cat(sd_treatment, sd_control, sd_P1, sd_P2)
# Plot group
df_2c <- data.frame(values=Y_2c) %>%
  mutate(groups=if_else(group_AB, "AB", "BA"),
         treatment=if_else(group_treatment, "Treatment", "Control"),
         period=if_else(group_period1, "P1", "P2"),
         eachfour=c(rep("AB_P1", n), rep("BA_P1", n),
                    rep("AB_P2", n), rep("BA_P2", n))
  )
ggplot(df_2c) +
  geom_density(aes(x=values, fill=groups), alpha=0.4) +
  theme_classic()
ggplot(df_2c) +
  geom_density(aes(x=values, fill=treatment), alpha=0.4) +
  theme_classic()
ggplot(df_2c) +
  geom_density(aes(x=values, fill=period), alpha=0.4) +
  theme_classic()
ggplot(df_2c) +
  geom_density(aes(x=values, fill=eachfour), alpha=0.4) +
  theme_classic()
```




## 2.D Yes treatment and period effects
```{r}
set.seed(123)
# Generate data
Y_2d <- generate.samples.crossover(
  n_total=n_total, b0=120, sigma_u=20, rho=0.75,
  treatment_effect=0.4, period_effect=0.03
  )
# Basic summary
# Means
mean_2d <- mean(Y_2d)
mean_AB <- mean(Y_2d[group_AB])
mean_BA <- mean(Y_2d[group_BA])
mean_treatment <- mean(Y_2d[group_treatment])
mean_control <- mean(Y_2d[group_control])
mean_P1 <- mean(Y_2d[group_period1])
mean_P2 <- mean(Y_2d[group_period2])
# Sds
sd_2d <- sd(Y_2d)
sd_within <- sqrt(get.var.within.crossover(Y_2d))
sd_AB <- sd(Y_2d[group_AB])
sd_BA <- sd(Y_2d[group_BA])
sd_treatment <- sd(Y_2d[group_treatment])
sd_control <- sd(Y_2d[group_control])
sd_P1 <- sd(Y_2d[group_period1])
sd_P2 <- sd(Y_2d[group_period2])
# Print
cat(mean_2d, mean_AB, mean_BA)
cat(mean_treatment, mean_control, mean_P1, mean_P2)
cat(sd_2d, sd_AB, sd_BA, sd_within)
cat(sd_treatment, sd_control, sd_P1, sd_P2)
# Plot group
df_2d <- data.frame(values=Y_2d) %>%
  mutate(groups=if_else(group_AB, "AB", "BA"),
         treatment=if_else(group_treatment, "Treatment", "Control"),
         period=if_else(group_period1, "P1", "P2"),
         eachfour=c(rep("AB_P1", n), rep("BA_P1", n),
                    rep("AB_P2", n), rep("BA_P2", n))
  )
ggplot(df_2d) +
  geom_density(aes(x=values, fill=groups), alpha=0.4) +
  theme_classic()
ggplot(df_2d) +
  geom_density(aes(x=values, fill=treatment), alpha=0.4) +
  theme_classic()
ggplot(df_2d) +
  geom_density(aes(x=values, fill=period), alpha=0.4) +
  theme_classic()
ggplot(df_2d) +
  geom_density(aes(x=values, fill=eachfour), alpha=0.4) +
  theme_classic()
```



## 2.E With random effect
```{r}
set.seed(123)
# Generate data
Y_2e <- generate.samples.crossover(
  n_total=n_total, b0=120, sigma_u=20, rho=0.75,
  treatment_effect=0.4, period_effect=0.03,
  do.random = TRUE
  )
# Basic summary
# Means
mean_2e <- mean(Y_2e)
mean_AB <- mean(Y_2e[group_AB])
mean_BA <- mean(Y_2e[group_BA])
mean_treatment <- mean(Y_2e[group_treatment])
mean_control <- mean(Y_2e[group_control])
mean_P1 <- mean(Y_2e[group_period1])
mean_P2 <- mean(Y_2e[group_period2])
# Sds
sd_2e <- sd(Y_2e)
sd_within <- sqrt(get.var.within.crossover(Y_2e))
sd_AB <- sd(Y_2e[group_AB])
sd_BA <- sd(Y_2e[group_BA])
sd_treatment <- sd(Y_2e[group_treatment])
sd_control <- sd(Y_2e[group_control])
sd_P1 <- sd(Y_2e[group_period1])
sd_P2 <- sd(Y_2e[group_period2])
# Print
cat(mean_2e, mean_AB, mean_BA)
cat(mean_treatment, mean_control, mean_P1, mean_P2)
cat(sd_2e, sd_AB, sd_BA, sd_within)
cat(sd_treatment, sd_control, sd_P1, sd_P2)
# Plot group
df_2e <- data.frame(values=Y_2e) %>%
  mutate(groups=if_else(group_AB, "AB", "BA"),
         treatment=if_else(group_treatment, "Treatment", "Control"),
         period=if_else(group_period1, "P1", "P2"),
         eachfour=c(rep("AB_P1", n), rep("BA_P1", n),
                    rep("AB_P2", n), rep("BA_P2", n))
  )
ggplot(df_2e) +
  geom_density(aes(x=values, fill=groups), alpha=0.4) +
  theme_classic()
ggplot(df_2e) +
  geom_density(aes(x=values, fill=treatment), alpha=0.4) +
  theme_classic()
ggplot(df_2e) +
  geom_density(aes(x=values, fill=period), alpha=0.4) +
  theme_classic()
ggplot(df_2e) +
  geom_density(aes(x=values, fill=eachfour), alpha=0.4) +
  theme_classic()
```


# Save datasets
```{r}
write.csv(Y_1a, "../data/parallel.csv")
write.csv(Y_1b, "../data/parallel_treatment.csv")
write.csv(Y_2a, "../data/crossover.csv")
write.csv(Y_2b, "../data/crossover_treatment.csv")
write.csv(Y_2c, "../data/crossover_period.csv")
write.csv(Y_2d, "../data/crossover_treatment_period.csv")
write.csv(Y_2e, "../data/crossover_treatment_period_random.csv")
```
















